Flowchart of Current Algorithm

### 2009-10-19 docs/flowchart/flowchart.tex
updates

\renewcommand{\arraystretch}{1.0}
There are some great charts in the appendix of the multilevel paper demonstrating what 
the superscript notation represents.
\section{Shorthand Functions}
See the multilevel paper for the latest.

@@

We now describe the full time advancement algorithm, making frequent
use of the shorthand developed above.  Here, we assume that the
problem is already properly initialized.  We describe the details of
the initialization in \S \ref{Sec:Initialization}.
The advance of the state through a single timestep appears as:

%--------------------------------------------------------------------------
% STEP 1
%--------------------------------------------------------------------------

\begin{description}

\item[Step 1.] {\em Define the provisional time-centered expansion, $S^{n+\myhalf,\star}$, 
provisional base state velocity, $w_0^{n+\myhalf,\star}$, and provisional base state velocity 
forcing.}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item At the beginning of each time step, we need an estimate for the time-centered
source term in the velocity divergence constraint, as given in paper III equation (19),
\begin{equation}
  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext 
  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip .
\label{eq:defineS} 
\end{equation}

If this is the first step of the calculation ($n=0$), we set
\begin{equation}
S^{\myhalf,\star} = \frac{S^0 + S^1}{2} \enskip ,
\end{equation}
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(BDRWZ), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}
\item Compute
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
\end{equation}
\item Compute $w_0^{n+\myhalf,\star}$:
\begin{description}
\item For planar geometry,\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,
\psi^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.

If we are using the volume discrepancy term, compute 
$\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
\item For spherical geometry,\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
\rhozero^n,p_0^n,g^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
\end{description}

\item Define the base state velocity forcing using equation (38) from paper III,
\begin{equation}
-\frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} 
= \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} 
\enskip, \label{eq:pizero}
\end{equation}
with the following discretization:
\begin{equation}
\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^{n,\star} = 
-\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} 
- w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star} \enskip ,
\end{equation} 
where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are defined through 
paper III equation (53),
\begin{equation}
w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}} 
\enskip,
\end{equation}
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = 
\frac{1}{\dt^n+\dt^{n-1} } 
\left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] 
\enskip.
\end{equation}
If $n=0$, we use $\dt^{-1} = \dt^0$.

%--------------------------------------------------------------------------
% STEP 2
%--------------------------------------------------------------------------
\item[Step 2.] {\em Construct the provisional time-centered advective velocity on 
edges, $\uadvone$.}

The local velocity field is described by paper III equation (37),
\begin{equation}
\frac{\partial\Ubt}{\partial t} = 
- \left(\Ubt+w_0\right) \cdot \nabla \Ubt
- \left(\Ubt \cdot \eb_r\right) \frac{\partial w_0}{\partial r} \eb_r
- \frac{1}{\rho} \nabla\pi
+ \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \eb_r
- \frac{(\rho-\rhozero)}{\rho} \; g \; \eb_r  \label{eq:utildeupd}  \enskip .
\end{equation}

We construct time-centered velocities at edges, $\uadvone$, 
using a second-order unsplit Godunov procedure followed by an elliptic
solve, as described in Appendix B of paper III.  We note that 
$\uadvone$ satisfies the discrete versions of paper III equation (21),
\begin{equation}
\int_{\Omega_H} \Ubt \cdot \eb_r \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
and paper III equation (39),
\begin{equation}
\nabla \cdot (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
\label{eq:tildeconstraint}
\end{equation}
with the discretization
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) \enskip .
\end{equation}
The discretization with the volume discrepancy correction is
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right]\right\} \enskip .
\end{equation}

%--------------------------------------------------------------------------
% STEP 3
%--------------------------------------------------------------------------
\item[Step 3.] {\em React the full state through the first time interval of $\dt / 2.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n] \rightarrow
[ \rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)} ]$.

\item Set $(\rho h)_0^n$ to be the lateral average of $(\rho h)^{(1)}$:
\begin{description}
\item For planar geometry, set $(\rho h)_0^n = $ {\bf Avg}$[(\rho h)^{(1)}]$
\item For spherical geometry, set $(\rho h)_0^n = (\rho h)_0^n +$ {\bf Avg}$[(\rho h)^{(1)} - (\rho h)_0^n]$
\end{description}

%--------------------------------------------------------------------------
% STEP 4
%--------------------------------------------------------------------------
\item[Step 4.] {\em Advect the base state and full state, through a time interval of $\dt.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf Advect Base Density}$[\rhozero^{n},w_0^{\nph,\star}] \rightarrow$ 
$[\rho_0^{(2a),\star}, \rho_0^{\nph,\star,{\rm pred}}]$.

\item \label{step:species update} 
Update the species.  Here we consider only the advection terms, 
neglecting the reaction terms, and solve a discretized version of
\begin{equation}
\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\Ub \rho X_k) = 0 \enskip .
\label{eq:species}
\end{equation}
The update consists of two steps:

  \begin{enumerate}
  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}

  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
  for the conservative update of $(\rho X_k)^{(1)}$. 
  Here we predict $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$ and 
  $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to obtain 
  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
  unsplit Godunov procedure, as described in paper II,
  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \eb_r$.  Convert the 
  perturbational edge state to a full state using
\begin{equation}
\rho^{(1),\nph,\star} = 
\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip .
\label{eq:rhoXedgestate}
\end{equation}
  In planar geometry, we use 4th-order spatial averaging to compute $\rho_0$ on 
  radial faces.  In spherical geometry, we first map $\rho_0$ to Cartesian cell-centers 
  using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.

  \item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$
  without explicitly including the reaction terms,
\begin{equation}
(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)} 
 - \dt \; \left\{ \nabla \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right)  
  (\rho X_k)^{(1),\nph,\star} \right] \right\} \enskip ,
\end{equation}
\begin{equation}
\rho^{(2),\star} = \sum_k (\rho X_k)^{(2),\star} \enskip ,
\end{equation}
\begin{equation}
X_k^{(2),\star} = (\rho X_k)^{(2),\star} / \rho^{(2),\star}
\end{equation}

\item Define an edge-centered $\etarho^{\nph,\star}$:

\begin{description}
\item For planar geometry,
\begin{equation}
 \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdot \eb_r + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
\end{equation}
\item For spherical geometry, first construct 
$\eta_{\rho}^{{\rm cart},\nph,\star} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf,\star}$ using:
\begin{equation}
\eta_{\rho}^{{\rm cart},\nph,\star} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvone \cdot \eb_d\right)n_d.
\end{equation}
Then, $\etarho^{\nph,\star}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph,\star}$,
\begin{equation}
\etarho^{\nph,\star} = \overline{\eta_{\rho}^{{\rm cart},\nph,\star}}.
\end{equation}
This gives a cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
radial edges, average the two neighboring radial cell-centers.
\end{description}

\item Set $\rho_0^{n+1,\star}$ to be the lateral average of $\rho^{(2),\star}$:
\begin{description}
\item For planar geometry, set $\rho_0^{n+1,\star} =$ {\bf Avg}$(\rho^{(2),\star})$
\item For spherical geometry, set $\rho_0^{n+1,\star} = \rho_0^{(2a),\star} +$ {\bf Avg}$(\rho^{(2),\star} - \rho_0^{(2a),\star})$
\end{description}

\item {\bf Compute Gravity}$[\rho_0^{n+1,\star}] \rightarrow [g^{n+1,\star}]$.

%{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
%\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
%\item For spherical geometry,\\
%{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
%\overline{S^{\nph,\star}}, \rho^{(2),\star}, X_k^{(2),\star}]$\\
%$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%\rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
\item {\bf Enforce HSE}$[p_0^n,\rho_0^{n+1,\star},g^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.

\item Compute $\psi$.
\begin{description}
\item For planar geometry,
\begin{equation}
\psi_j^{n+\myhalf,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf,\star} 
+ \eta_{\rho,j+\myhalf}^{n+\myhalf,\star}\right) g \enskip .
\end{equation}
\item For spherical geometry, first compute:
\begin{equation}
\gammabar^{(1)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, 
X_k^{(1)}\right) \right] \enskip ,
\end{equation}
\begin{equation}
\gammabar^{(2),\star} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2),\star}, p_0^{n+1,\star}, 
X_k^{(2),\star}\right) \right] \enskip .
\end{equation}
Then, define a base time and time-centered $\psi$:
\begin{equation}
\psi_j^{n,\star} = 
\left(\gammabar^{(1)} p_0^n\right)_j
\left \{ \Sbar_j^{n+\myhalf,\star} - 
\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip ,
\end{equation}
\begin{eqnarray}
\psi_j^{n+\myhalf,\star} &=& 
\left(\frac{\gammabar^{(1)}+\gammabar^{(2),\star}}{2}\right)_j
\left(\frac{p_0^n+p_0^{n+1,\star}}{2}\right)_j \nonumber \\
&& \left \{ \Sbar_j^{n+\myhalf,\star} - 
\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip .
\end{eqnarray}
\end{description}

\item Update $(\rho h)_0$.
\begin{description}
\item For planar geometry,\\
{\bf Advect Base Enthalpy Planar}$[(\rho h)_0^{n}, w_0^{\nph,\star}, \psi^{\nph,\star}]
\rightarrow [(\rho h)_0^{n+1,\star}]$.
\item For spherical geometry,\\
{\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^{n}, w_0^{\nph,\star}, 
\psi^{n,\star},\psi^{\nph,\star}] \rightarrow [(\rho h)_0^{n+1,\star}]$.
\end{description}

\item Update the enthalpy.  Here we consider only the advective terms, 
neglecting the reaction terms and thermal diffusion terms, and solve a discretized version of
  \begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h)
+ \psi + (\Ubt \cdot \eb_r) \frac{\partial p_0}{\partial r} 
\enskip , \label{eq:rhohupdadv} 
  \end{equation}
For spherical geometry, experience has shown that solving the analytically equivalent form,
\begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h)
+ \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt  
\enskip , \label{eq:rhohupdadv2} 
\end{equation}
results in a better solution.

  \begin{enumerate}
  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}

  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
    for the conservative update of $(\rho h)^{(1)}.$  Here we predict 
    $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0$ to time-centered edges to obtain 
    $(\rho h)^{'(1),\nph,\star}$, 
    using a second-order unsplit Godunov procedure, 
    as described in paper II, Appendix A, using $\Vb =
    \uadvone+w_0^{\nph,\star} \eb_r$.  We do not include the reaction
    terms in the enthalpy prediction, since we accounted for them
    already in {\bf React State}.  Convert the perturbational edge
    state to a full state using

\begin{equation}
(\rho h)^{(1),\nph,\star} = 
(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}
\enskip .
\end{equation}
  For planar geometry, we use 4th-order spatial averaging to compute $(\rho h)_0$
  on radial faces.  In spherical geometry, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
  and $\rho_0$ to Cartesian cell-centers using the mapping in section 
  \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, use 2nd-order spatial 
  averaging to compute $h_0$ and $\rho_0$ on all faces, and then multiply these 
  terms to get $(\rho h)_0$ on all faces.

  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without
  explicitly including the reaction terms.

\begin{description}
\item For planar geometry,

  \begin{eqnarray}
  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla
      \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho
      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ && + \dt \;
    \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial
      p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nph,\star} \enskip ,
  \end{eqnarray}

\item For spherical geometry,

  \begin{eqnarray}
  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla
      \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho
      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ 
    && + \dt \; \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) 
       - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
    && + \dt \; \psi^{\nph,\star} \enskip ,
  \end{eqnarray}

\end{description}

If {\tt do\_eos\_h\_above\_cutoff=T} (which is the default setting), then if
$\rho^{(2),\star} < {\tt base\_cutoff\_density}$, then we recompute enthalpy using
\begin{equation}
(\rho h)^{(2),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
\end{equation}

%--------------------------------------------------------------------------
% STEP 4.1
%--------------------------------------------------------------------------

\item If we are using thermal diffusion, diffuse the enthalpy through a time interval of 
$\dt$.  First, define $(\rho h)^{(1a),\star} = (\rho h)^{(2),\star}$.  We recompute
$(\rho h)^{(2),\star}$ to account for thermal diffusion.  Here we begin
with the enthalpy equation, but consider only the 
diffusion terms,
\begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = 
 \nabla\cdot\kth\nabla T 
\enskip . \label{eq:rhohupdthermal} 
  \end{equation}
We can recast this as an enthalpy-diffusion equation by applying the
chain-rule to $h(p_0,T,X_k)$,
\begin{equation}
\nabla h = h_p \nabla p_0 + c_p \nabla T + \sum_k \xi_k \nabla X_k \enskip ,
\end{equation}
giving
\begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = 
 \nabla\cdot \frac{\kth}{c_p}\nabla h -  
 \sum_k \nabla\cdot \frac{\xi_k \kth}{c_p}\nabla X_k -
 \nabla\cdot \frac{h_p \kth}{c_p}\nabla p_0 
\enskip . \label{eq:rhohupdthermal2} 
  \end{equation}

Compute $\kth^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.  The update is given by
\begin{eqnarray}
(\rho h)^{(2),\star} &=& (\rho h)^{(1a),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n+1,\star} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n}\right),
\end{eqnarray}
which is numerically implemented as a diffusion equation for $h^{(2),\star}$,
\begin{eqnarray}
\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(1a),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n+1,\star} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n}\right),
\end{eqnarray}
\item Update the temperature using the equation of state.  For planar geometry,
we update using
\begin{equation}
T^{(2),\star} = T\left(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star}\right) \enskip ,
\end{equation}
but in spherical geometry, we set {\tt use\_tfromp = T}, and we use
\begin{equation}
T^{(2),\star} = T\left(\rho^{(2),\star}, p_0^{n+1,\star}, X_k^{(2),\star}\right) \enskip .
\end{equation}
The latter has the effect of completely decoupling the enthalpy equation from the 
evolution.

%--------------------------------------------------------------------------
% STEP 5
%--------------------------------------------------------------------------
\item[Step 5.] {\em React the full state through a second time interval of $\dt / 2.$}

{\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star}, 
(\rho\Hext)^{(2),\star}, p_0^{n+1,\star}] $\\
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, 
X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$

%--------------------------------------------------------------------------
% STEP 6
%--------------------------------------------------------------------------
\item[Step 6.] {\em Define the time-centered expansion, $S^{n+\myhalf}$, base state
velocity, $w_0^{n+\myhalf}$, and base state velocity forcing.}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}
\item Define
\begin{equation}
  S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + 
  \sigma \Hnuc^{(2),\star} +
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star}  
  + \sigma \Hext + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T^{n+1,\star} \enskip ,
\end{equation} 
where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} / \rho^{(2),\star}$
and the thermodynamic quantities are defined using $\rho^{n+1,\star}, X_k^{n+1,\star},$ 
and $T^{n+1,\star}$ as inputs to the equation of state.
Then define
\begin{equation}
 S^\nph = \frac{S^n + S^{n+1,\star}}{2} \enskip. 
\end{equation}

\item Compute
\begin{equation}
\overline{S^{\nph}} = {\mathrm{\bf Avg}} (S^{\nph}) \enskip.
\end{equation}

>>>>

\renewcommand{\arraystretch}{1.0}
There are some great charts in the appendix of the multilevel paper demonstrating what 
the superscript notation represents.

@@

Here are changes to the algorithm from the multilevel paper description.
The initialization has also been included with more detail than was given in paper III.
\subsection{Volume Discrepancy Changes}
\item In {\bf Step 2B}, to compute $w_0$, we need to account for the volume discrepancy
term by computing $p_{\rm EOS} = \overline{p(\rho,h,X_k)^n}$.
\item In {\bf Step 3}, the MAC projection should account for the volume discrepancy term:
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left\{ \left(S^{\nph,\star\star} - \overline{S^{\nph,\star\star}}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right]\right\}.
\end{equation}
\item In {\bf Step 6B}, to compute $w_0$, we need to account for the volume discrepancy
term by computing $p_{\rm EOS} = \left[\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}\right]/2$.
\item In {\bf Step 7}, the MAC projection should account for the volume discrepancy term:
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star}\left\{\left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) + \frac{f}{\overline{\Gamma_1^{\nph,\star}} p_0^{\nph,\star}} \left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\},
\end{equation}
where $p(\rho,h,X_k)^{\nph,\star} = \left[p(\rho,h,X_k)^n +p(\rho,h,X_k)^{n+1,\star}\right]/2$.
\item In {\bf Step 11}, the approximate projection should account for the volume
discrepancy term:
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S^{n+1}} \right)
+ \frac{f}{\overline{\Gamma_1^{n+1}} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\}.
\end{equation}
\subsection{Thermal Diffusion Changes}
Immediately after {\bf Step 4H}, diffuse the enthalpy through 
a time interval of $\dt$.  First, define $(\rho h)^{(1a),\star} = (\rho h)^{(2),\star}$.  
We recompute $(\rho h)^{(2),\star}$ to account for thermal diffusion.  Here we begin
with the enthalpy equation, but consider only the 
diffusion terms,
\begin{equation}
\frac{\partial (\rho h)}{\partial t} = \nabla\cdot\kth\nabla T.
  \end{equation}
We can recast this as an enthalpy-diffusion equation by applying the
chain-rule to $h(p_0,T,X_k)$,
\begin{equation}
\nabla h = h_p \nabla p_0 + c_p \nabla T + \sum_k \xi_k \nabla X_k \enskip ,
\end{equation}
giving
\begin{equation}
  \frac{\partial (\rho h)}{\partial t}  = 
 \nabla\cdot \frac{\kth}{c_p}\nabla h -  
 \sum_k \nabla\cdot \frac{\xi_k \kth}{c_p}\nabla X_k -
 \nabla\cdot \frac{h_p \kth}{c_p}\nabla p_0.
  \end{equation}
Compute $\kth^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.  The update is given by
\begin{eqnarray}
(\rho h)^{(2),\star} &=& (\rho h)^{(1a),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n+1,\star} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n}\right),
\end{eqnarray}
which is numerically implemented as a diffusion equation for $h^{(2),\star}$,
\begin{eqnarray}
\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(1a),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n+1,\star} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n}\right),
\end{eqnarray}

Immediately after {\bf Step 8H}, diffuse the enthalpy through a time interval of 
$\dt$.  First, define $(\rho h)^{(1a)} = (\rho h)^{(2)}$.  We recompute $(\rho h)^{(2)}$ to 
account for thermal diffusion.  Compute $\kth^{(2),\star}, c_p^{(2),\star}$, and 
$\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to 
the equation of state.  The update is given by
\begin{eqnarray}
(\rho h)^{(2)} &=& (\rho h)^{(1a)} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{n+1} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n}\right),
\end{eqnarray}
which is numerically implemented as a diffusion equation for $h^{(2)}$.
\begin{eqnarray}
\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(1a)} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{n+1} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n}\right),
\end{eqnarray}
